---
title: "Smoking Status and Blood Pressure"
author: "Nontu"
output: html_document
---

# Introduction

High blood pressure is an important public health concern. This mini report explores a small example dataset to compare blood pressure levels between smokers and non-smokers.

# Objective

The objective of this analysis is to describe blood pressure patterns by smoking status.

# Methods

A small example dataset was created in R. The dataset includes patient age, smoking status, and blood pressure.

```{r}
patient_id <- 1:8
age <- c(25, 40, 35, 60, 50, 28, 45, 55)
smoker <- c("Yes", "No", "Yes", "No", "Yes", "No", "Yes", "No")
blood_pressure <- c(120, 140, 135, 150, 145, 125, 155, 138)

patients <- data.frame(patient_id, age, smoker, blood_pressure)

patients
```

# Results

## Average blood pressure

```{r}
mean(blood_pressure)
```

## Average blood pressure by smoking status

```{r}
tapply(blood_pressure, smoker, mean)
```

## Create a risk category

```{r}
patients$risk <- ifelse(blood_pressure > 140, "High", "Normal")

patients
```

## Patients with high blood pressure

```{r}
patients[patients$risk == "High", ]
```

# Interpretation

The analysis compares blood pressure values between smokers and non-smokers in this example dataset. The risk category identifies patients with blood pressure above 140 as high risk.

# Conclusion

This project demonstrates how R Markdown can be used to combine text, R code, results, and interpretation in one reproducible public health report.